Local Type Inference for Polarised System F with Existentials

### 2024-01-23 oopsla24/tex/04-algorithm.tex
wip

@@

\paragraph{Algorithmic Contexts and Well-formedness}
To specify when algorithmic types are well-formed, 
we define algorithmic contexts $[[Ξ]]$ as sets of algorithmic variables.
Then $[[Γ ; Ξ ⊢ uP]]$ and $[[Γ ; Ξ ⊢ uN]]$ represent the well-formedness
judgment of algorithmic terms defined as expected. 
Informally, they check that all free declarative variables are in $[[Γ]]$, 
and all free algorithmic variables are in $[[Ξ]]$.
In addition to the rules repeating the declarative definition \cref{todo}, 
we have two base cases for the algorithmic variables:
 \ruleref{\ottdruleWFATPUVarLabel} and \ruleref{\ottdruleWFATNUVarLabel} (see \cref{fig:algo-wf})

\begin{figure}[h]
\begin{multicols}{2}
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleWFATPUVarLabeled{}}
  \columnbreak\\
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleWFATNUVarLabeled{}}
\end{multicols}
\label{fig:algo-wf}
\caption{Well-formedness of Algorithmic Types}
\end{figure}

@@

    % Define nodes
    \node[] (1) {Negative Subtyping\\$[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$\\(\cref{sec:subtyping})};
    \node[below of=1] (3) {Positive Subtyping\\$[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$\\(\cref{sec:subtyping})};
    \node[left=1.5cm of 3] (2) {Constraint Merge\\$[[Θ ⊢ SC1 & SC2 = SC3]]$\\(\cref{sec:constraint-merge})};
    \node[right=1.5cm of 3] (5) {Unification\\ $[[Γ ; Θ ⊨ uN ≈u iM ⫤ UC]]$\\ $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ UC]]$\\(\cref{sec:unification})};
    \node[below of=3] (4) {Upgrade\\$[[upgrade Γ ⊢ iP to Δ = iQ]]$\\(\cref{sec:lub})};
    \node[below of=2] (6) {Least Upper Bound\\$[[Γ ⊨ iP1 ∨ iP2 = iQ]]$\\(\cref{sec:lub})};
    \node[below of=6] (7) {Anti-Unification\\$[[Γ ⊨ iP1 ≈au iP2 ⫤ ( Ξ , uQ , aus1 , aus2 )]]$\\$[[Γ ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )]]$\\(\cref{sec:antiunification})};
    \node[below of=5] (8) {Unification Constraint Merge\\$[[Θ ⊢ UC1 & UC2 = UC3]]$\\(\cref{sec:constraint-merge})};

@@

\paragraph{Shifts} Rules \ruleref{\ottdruleAShiftDLabel} and
\ruleref{\ottdruleAShiftULabel} cover the downshift and the upshift cases,
respectively. If the input types are constructed by shifts, then the subtyping
can only hold if they are equivalent. This way, the algorithm must find the
instantiations of the algorithmic variables on the left-hand side, which make it
equivalent to the right-hand side. For this purpose, the algorithm invokes the
unification procedure \cref{todo} preceded by normalization of the input types.
It returns the resulting constraints given by the unification algorithm.

\paragraph{Quantifiers}  
Rules \ruleref{\ottdruleAForallLabel} and 
\ruleref{\ottdruleAExistsLabel} are symmetric. 
Declaratively, the quantified variables on the left-hand side must 
be instantiated with types, which are not known beforehand.
We address this problem by algorithmization (\cref{todo})
of the quantified variables.
The rule introduces fresh algorithmic variables
$[[puas]]$ or $[[nuas]]$,
puts them into the constraint context $[[Θ]]$
(specifying that they must be instantiated in the extended context
$[[Γ, pbs]]$ or $[[Γ, nbs]]$) and substitute the quantified variables
for them in the input type.

@@

\paragraph{Functions}
To infer the subtyping of the function types, the algorithm
makes two calls: 
\begin{enumerate*}
  \item[(i)] a recursive call ensuring the subtyping of the result types, and
  \item[(ii)] a call to positive subtyping (or rather super-typing) on the argument types.
\end{enumerate*}
The resulting constraints are merged (using a special procedure defined later in \cref{todo})
and returned as the output.

@@

The potential problem here is that the type $[[iP]]$
might be not well-formed in the context required for $[[α̂⁺]]$ by $[[Θ]]$, 
because this context might be smaller than the current context $[[Γ]]$.
As we wish the resulting constraint set to be sound \wrt $[[Θ]]$,
we cannot simply put $[[α̂⁺ :≥ iP]]$ into the output. 
Prior to that, we update the type $[[iP]]$ to its lowest supertype $[[iQ]]$
well-formed in $[[Θ(α̂⁺)]]$. It is done by the \emph{upgrade} procedure,
which we discuss in detail in \cref{todo}.

@@

Since only the left-hand side may contain algorithmic variables,
that the unification instantiates, we could have called this procedure \emph{matching}.
However, in \cref{todo} we will discuss several modifications of the 
type system, where this invariant is not preserved, and thus, this procedure
becomes a genuine first-order pattern unification \cite{miller1991:unification}.

@@

\paragraph{Functions}
  Rule \ruleref{\ottdruleUArrowLabel} unifies two functional types. 
  First, it unifies the argument types and their result types recursively
  Then it merges the resulting constraints using the constraint merge procedure (\cref{todo}).

  Notice that the resulting constraints can only have \emph{unification} entries.
  It means that they can be merged in a simpler way than general constraints.
  In particular, the merging procedure does not call any of the subroutines discussed here,
  but rather simply checks the matching constraint entries for equality.

@@

If both input entries are supertype constraints: $[[α̂⁺ :≥ iP]]$ and $[[α̂⁺ :≥ iQ]]$,
then their conjunction is $[[α̂⁺ :≥ iP ∨ iQ]]$, as given by \ruleref{\ottdruleSCMESupSupLabel}.
The least upper bound---$[[iP ∨ iQ]]$ is the least supertype of both $[[iP]]$ and $[[iQ]]$,
and this way, $[[α̂⁺ :≥ iP ∨ iQ]]$ is the weakest constraint entry that implies
$[[α̂⁺ :≥ iP]]$ and $[[α̂⁺ :≥ iQ]]$. The algorithm for finding the least upper bound
is discussed in \cref{todo}.

@@

The anti-unification algorithm \cite{todo}, 
is a procedure dual to unification
For two given (potentially different) expressions, 
it finds the most specific generalizer---the most detailed pattern that 
matches both of the input expressions. As evidence, it can also return 
two substitutions that instantiate the `holes' of the pattern to the input expressions.

@@

At the high level, the algorithm scheme follows
the standard approach \cite{todo} consisting of two principles:
\begin{enumerate}
  \item[(i)] if the input terms start with the same constructor,
    we anti-unify the corresponding parts recursively and 
    unite the results. This principle is followed by 
    all the rules except \ruleref{\ottdruleAUAULabel},
    which works as follows:
  \item[(ii)] if the first principle does not apply to the input terms $[[iN]]$
    and $[[iM]]$ (for instance, if they have different outer constructors), the
    anti-unification algorithm returns a `hole' such that one substitution maps
    it to $[[iN]]$ and the other maps it to $[[iM]]$. The name of this `hole'
    should have a name uniquely defined by the pair $([[iN]], [[iM]])$, so
    that it automatically merges with other `holes' mapped to the same
    pair of types, and thus, the initiality of the generalizer is ensured. 
\end{enumerate}

@@

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.2cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]
    % Define nodes
    \node[] (1) {Positive Inference\\$[[Γ; Φ ⊨ v : iP]]$};
    \node[right=1.4cm of 1] (2) {Negative Inference\\$[[Γ; Φ ⊨ c : iN]]$};
    \node[right=1.4cm of 2] (3) {Application Inference\\$[[Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC]]$};
    \node[below of=2,xshift=+2cm] (4) {Constraint Merge\\$[[Θ ⊢ SC1 & SC2 = SC3]]$\\(\cref{sec:constraint-merge})};
    \node[below of=4] (5) {Positive Subtyping\\$[[Γ ; Θ ⊨ uP1 ≥ iP2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \node[left=0.75cm of 4] (6) {Negative Subtyping\\$[[Γ ; Θ ⊨ uN1 ≤ iN2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \draw[->] ([yshift=-.2cm]2.west) to ([yshift=-.2cm]1.east);
    \draw[->] ([yshift=+.2cm]1.east) to ([yshift=+.2cm]2.west);
    \draw[->, bend right=20] (3) to (1);
    \draw[->] (2) to (3);
    \draw[->] (2) to (4);
    \draw[->] (3) to (4);
    \draw[->, bend right] (1.south west) to (5);
    \draw[->, bend left] (3) to (5);
    \draw[->] (2) to (6);
  \end{tikzpicture}  
  \caption{Dependency graph of the typing algorithm}
  \label{fig:alg-typing-graph}
\end{figure}

@@

>>>>

\label{sec:algorithm}

@@

\paragraph{Algorithmic Contexts and Well-formedness}
To specify when algorithmic types are well-formed, 
we define algorithmic contexts $[[Ξ]]$ as sets of algorithmic variables.
Then $[[Γ ; Ξ ⊢ uP]]$ and $[[Γ ; Ξ ⊢ uN]]$ represent the well-formedness
judgment of algorithmic terms defined as expected. 
Informally, they check that all free declarative variables are in $[[Γ]]$, 
and all free algorithmic variables are in $[[Ξ]]$.
Most of the rules are inherited from the well-formedness of 
\emph{declarative} types: the declarative variables are
checked to belong to the context $[[Γ]]$, the quantifiers extend the context,
type constructors are well-formed congruently.
As we extend the syntax with algorithmic variables, we also add 
two base-case rules for them:
 \ruleref{\ottdruleWFATPUVarLabel} and \ruleref{\ottdruleWFATNUVarLabel} 
 (see \cref{fig:algo-wf}).

\begin{figure}[h]
\begin{multicols}{2}
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleWFATPUVarLabeled{}}
  \columnbreak\\
  $\vcenter{\hbox{\vdots}}$\\
  \ottusedrule{\ottdruleWFATNUVarLabeled{}}
\end{multicols}
\caption{Well-formedness of Algorithmic Types}
\label{fig:algo-wf}
\end{figure}

@@

    % Define nodes
    \node[] (1) {Negative Subtyping\\$[[Γ ; Θ ⊨ uN ≤ iM ⫤ SC]]$\\(\cref{fig:alg-subtyping})};
    \node[below of=1] (3) {Positive Subtyping\\$[[Γ ; Θ ⊨ uP ≥ iQ ⫤ SC]]$\\(\cref{fig:alg-subtyping})};
    \node[left=1.5cm of 3] (2) {Constraint Merge\\$[[Θ ⊢ SC1 & SC2 = SC3]]$\\(\cref{sec:constraint-merge})};
    \node[right=1.5cm of 3] (5) {Unification\\ $[[Γ ; Θ ⊨ uN ≈u iM ⫤ UC]]$\\ $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ UC]]$\\(\cref{sec:unification})};
    \node[below of=3] (4) {Upgrade\\$[[upgrade Γ ⊢ iP to Δ = iQ]]$\\(\cref{sec:lub})};
    \node[below of=2] (6) {Least Upper Bound\\$[[Γ ⊨ iP1 ∨ iP2 = iQ]]$\\(\cref{sec:lub})};
    \node[below of=6] (7) {Anti-Unification\\$[[Γ ⊨ iP1 ≈au iP2 ⫤ ( Ξ , uQ , aus1 , aus2 )]]$\\$[[Γ ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )]]$\\(\cref{sec:antiunification})};
    \node[below of=5] (8) {Unification Constraint Merge\\$[[Θ ⊢ UC1 & UC2 = UC3]]$\\(\cref{sec:constraint-merge})};

@@

\paragraph{Shifts} Rules \ruleref{\ottdruleAShiftDLabel} and
\ruleref{\ottdruleAShiftULabel} cover the downshift and the upshift cases,
respectively. If the input types are constructed by shifts, then the subtyping
can only hold if they are equivalent. This way, the algorithm must find the
instantiations of the algorithmic variables on the left-hand side
such that these instantiations make the left-hand side and the ride-hand side
equivalent. For this purpose, the algorithm invokes the
unification procedure (\cref{sec:unification}) preceded by the normalization of the input types.
It returns the resulting constraints given by the unification algorithm.

\paragraph{Quantifiers}  
Rules \ruleref{\ottdruleAForallLabel} and 
\ruleref{\ottdruleAExistsLabel} are symmetric. 
Declaratively, the quantified variables on the left-hand side must 
be instantiated with types, which are not known beforehand.
We address this problem by algorithmization 
of the quantified variables (see \cref{sec:algo-syntax}).
The rule introduces fresh algorithmic variables
$[[puas]]$ or $[[nuas]]$,
puts them into the constraint context $[[Θ]]$
(specifying that they must be instantiated in the extended context
$[[Γ, pbs]]$ or $[[Γ, nbs]]$) and substitute the quantified variables
for them in the input type.

@@

\paragraph{Functions}
To infer the subtyping of the function types, the algorithm
makes two calls: 
\begin{enumerate*}
  \item[(i)] a recursive call ensuring the subtyping of the result types, and
  \item[(ii)] a call to positive subtyping (or rather super-typing) on the argument types.
\end{enumerate*}
The resulting constraints are merged
(using a special procedure defined in \cref{sec:constraint-merge})
and returned as the output.

@@

The potential problem here is that the type $[[iP]]$
might be not well-formed in the context required for $[[α̂⁺]]$ by $[[Θ]]$, 
because this context might be smaller than the current context $[[Γ]]$.
As we wish the resulting constraint set to be sound \wrt $[[Θ]]$,
we cannot simply put $[[α̂⁺ :≥ iP]]$ into the output. 
Prior to that, we update the type $[[iP]]$ to its lowest supertype $[[iQ]]$
well-formed in $[[Θ(α̂⁺)]]$. It is done by the \emph{upgrade} procedure,
which we discuss in detail in \cref{sec:lub}.

@@

Since only the left-hand side may contain algorithmic variables,
that the unification instantiates, we could have called this procedure \emph{matching}.
However, in \cref{sec:weakening-invariant}, we will discuss several modifications of the 
type system, where this invariant is not preserved, and therefore, this procedure
requires general first-order pattern unification \cite{miller1991:unification}.

@@

\paragraph{Functions}
  Rule \ruleref{\ottdruleUArrowLabel} unifies two functional types. 
  First, it unifies the argument types and their result types recursively.
  Then it merges the resulting constraints using the procedure described in \cref{sec:constraint-merge}.

  Notice that the resulting constraints can only have \emph{unification} entries.
  It means that they can be merged in a simpler way than general constraints.
  In particular, the merging procedure does not call any of the subtyping subroutines,
  but rather simply checks the matching constraint entries for equality.

@@

If both input entries are supertype constraints: $[[α̂⁺ :≥ iP]]$ and $[[α̂⁺ :≥ iQ]]$,
then their conjunction is $[[α̂⁺ :≥ iP ∨ iQ]]$, as given by \ruleref{\ottdruleSCMESupSupLabel}.
The least upper bound---$[[iP ∨ iQ]]$ is the least supertype of both $[[iP]]$ and $[[iQ]]$,
and this way, $[[α̂⁺ :≥ iP ∨ iQ]]$ is the weakest constraint entry that implies
$[[α̂⁺ :≥ iP]]$ and $[[α̂⁺ :≥ iQ]]$. The algorithm finding the least upper bound
is discussed in \cref{sec:lub}.

@@

The anti-unification algorithm \cite{plotkin1970:generalization,reynolds1970:transform}, 
is a procedure dual to unification
For two given (potentially different) expressions, 
it finds the most specific generalizer---the most detailed pattern that 
matches both of the input expressions. As evidence, it can also return 
two substitutions that instantiate the `holes' of the pattern to the input expressions.

@@

At the high level, the algorithm scheme follows
the standard approach \cite{plotkin1970:generalization}
represented as a recursive procedure. Specifically, it follows two principles:
\begin{enumerate}
  \item[(i)] if the input terms start with the same constructor,
    we anti-unify the corresponding parts recursively and 
    unite the results. This principle is followed by 
    all the rules except \ruleref{\ottdruleAUAULabel},
    which works as follows:
  \item[(ii)] if the first principle does not apply to the input terms $[[iN]]$
    and $[[iM]]$ (for instance, if they have different outer constructors), the
    anti-unification algorithm returns a `hole' such that one substitution maps
    it to $[[iN]]$ and the other maps it to $[[iM]]$. This `hole'
    should have a name uniquely defined by the pair $([[iN]], [[iM]])$, so
    that it automatically merges with other `holes' mapped to the same
    pair of types, and thus, the initiality of the generalizer is ensured. 
\end{enumerate}

@@

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    [>={Stealth[scale=2]},node distance=2.2cm,every node/.style={draw,rectangle},every text node part/.style={align=center}]
    % Define nodes
    \node[] (1) {Positive Inference\\$[[Γ; Φ ⊨ v : iP]]$};
    \node[right=1.2cm of 1] (2) {Negative Inference\\$[[Γ; Φ ⊨ c : iN]]$};
    \node[right=1.2cm of 2] (3) {Application Inference\\$[[Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC]]$};
    \node[below=0.8cm of 2] (7) {Constraint Singularity\\$[[SC singular with uσ]]$\\(\cref{sec:constraint-singularity})};
    \node[right=0.3cm of 7] (4) {Constraint Merge\\$[[Θ ⊢ SC1 & SC2 = SC3]]$\\(\cref{sec:constraint-merge})};
    \node[below=0.3cm of 7] (5) {Positive Subtyping\\$[[Γ ; Θ ⊨ uP1 ≥ iP2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \node[left=0.3cm of 7] (6) {Negative Subtyping\\$[[Γ ; Θ ⊨ uN1 ≤ iN2 ⫤ SC]]$\\(\cref{sec:subtyping-algorithm})};
    \draw[->] ([yshift=-.2cm]2.west) to ([yshift=-.2cm]1.east);
    \draw[->] ([yshift=+.2cm]1.east) to ([yshift=+.2cm]2.west);
    \draw[->, bend right=20] (3) to (1);
    \draw[->] (2) to (3);
    \draw[->] (2) to (4);
    \draw[->] (2) to (7);
    \draw[->] (3) to (4);
    \draw[->] (1.south west) |- (5);
    \draw[->] (3) |- (5);
    \draw[->] (2) to (6);
  \end{tikzpicture}  
  \caption{Dependency graph of the typing algorithm}
  \label{fig:alg-typing-graph}
\end{figure}

@@

\subsection{Constraint Singularity}
\label{sec:constraint-singularity}

The singularity algorithm is used to check that the inferred type of an
application is unique. In this case, the annotation in the applicative let
binder can be omitted. In the algorithmic system, the uniqueness means that the
\emph{constraint set} produced by the application inference has a single
substitution satisfying it.

To implement the singularity algorithm, we define a partial function $[[SC
singular with uσ]]$, taking a subtyping constraint $[[SC]]$ as an argument and
returning a substitution $[[uσ]]$---the only solution of $[[SC]]$. The
constraint $[[SC]]$ consists of constraint entries. Thus, we define the
singularity using the singularity of constraint entries: each entry of $[[SC]]$
must have a unique instantiation and the resulting substitution $[[uσ]]$ must be
the combination of these instantiations (see \cref{fig:constraint-singularity}).

\begin{figure}[h]
  For a constraint set $[[SC]]$, we write $[[SC singular with uσ]]$ if and only if:
  \begin{itemize}
    \item for any \emph{positive} constraint entry $[[scE ∊ SC]]$, there exists $[[iP]]$ such that $[[scE singular with iP]]$,
      and for the variable $[[β̂⁺]]$ restricted by $[[scE]]$, $[[ [uσ]β̂⁺ = iP ]]$;
    \item the symmetric property holds for all \emph{negative} $[[scE ∊ SC]]$; and
    \item for any $[[α̂± ∉ dom(SC)]]$, $[[ [uσ]α̂± = α̂± ]]$.
  \end{itemize}
  \caption{Constraint Singularity}
  \label{fig:constraint-singularity}
\end{figure}

The singularity of constraint \emph{entries} is defined in
\cref{fig:constraint-entry-singularity}. The \emph{equivalence} entries are
always singular, as the only possible type satisfying them is the one given in
the constraint itself, which is reflected in \ruleref{\ottdruleSINGNEqLabel} and
\ruleref{\ottdruleSINGPEqLabel}. The \emph{subtyping} constraints are trickier.
As will be discussed in \cref{sec:proof-lub-upgrade}, variables 
(and equivalent types such as $[[∃nas.pa]]$) do not have 
\emph{proper} supertypes, and thus, the constraints of a form 
$[[pua :≥ ∃nas.pa]]$ are singular with the only possible normalized solution $[[pa]]$
\ruleref{\ottdruleSINGSupVarLabel}. However, if the body of the existential type
is guarded by a shift $[[∃nas.↓iN]]$, it is singular if and only if $[[iN]]$ is
equivalent to some $[[αi⁻ ∊ {nas}]]$ bound by the quantifier, see the rule
\ruleref{\ottdruleSINGSupShiftLabel}.

\begin{figure}[h]
  \hfill\\
  \ottdefnSINGLabeled
  \caption{Singular Constraint Entry}
  \label{fig:constraint-entry-singularity}
\end{figure}

### 2024-01-23 oopsla24/tex/05-proof.tex
wip

\subsection{Least Upper Bound and Upgrade}

@@

\subsection{Subtyping}

    As for other properties, the correctness of subtyping means that
    the algorithm produces a valid result (soundness) 
    whenever it exists (completeness).
    In \cref{fig:subtyping-algo}, one can see that the positive and 
    negative subtyping relations are not \emph{mututally} recursive: 
    negative subtyping algorithm uses the positive subtyping, 
    but not vice versa. 
    Because of that, the inductive proofs of the \emph{positive} subtyping correctness are done independently.

@@

The soundness requires completeness in the case of
\ruleref{\ottdruleATAppLetLabel}. To prove the soundness, 
in other words, to transform this rule into its
declarative counterpart \ruleref{\ottdruleDTAppLetLabel}, one needs
to prove \emph{uniqueness} of the inferred application type $[[↑iQ]]$. 
To show the uniqueness, we need to conclude that any other \emph{declarative} tree
(which is given by the induction hypothesis) infers an equivalent type. 
The only way to do so is by applying the soundness of singularity (\cref{sec:singularity-proo}), however, first, the declarative tree must be converted to the algorithmic one (by completeness!).
This way, the soundness and completeness proofs are mutually dependent.

>>>>

\subsection{Least Upper Bound and Upgrade}
\label{sec:proof-lub-upgrade}

@@

\subsection{Subtyping}
    \label{sec:proof-subtyping}

    As for other properties, the correctness of subtyping means that
    the algorithm produces a valid result (soundness) 
    whenever it exists (completeness).
    In \cref{fig:alg-subtyping}, one can see that the positive and 
    negative subtyping relations are not \emph{mututally} recursive: 
    negative subtyping algorithm uses the positive subtyping, 
    but not vice versa. 
    Because of that, the inductive proofs of the \emph{positive} subtyping correctness are done independently.

@@

The soundness requires completeness in the case of
\ruleref{\ottdruleATAppLetLabel}. To prove the soundness, 
in other words, to transform this rule into its
declarative counterpart \ruleref{\ottdruleDTAppLetLabel}, one needs
to prove \emph{uniqueness} of the inferred application type $[[↑iQ]]$. 
To show the uniqueness, we need to conclude that any other \emph{declarative} tree
(which is given by the induction hypothesis) infers an equivalent type. 
The only way to do so is by applying the soundness of singularity (\cref{sec:singularity-proof}), 
however, first, the declarative tree must be converted to the algorithmic one (by completeness!).
This way, the soundness and completeness proofs are mutually dependent.
